The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba

### 2020-11-27 analysis/paper/paper.Rmd
missing final text and location

---
title: 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/template.docx' # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/journal-of-archaeological-science.csl' # Insert path for the bib-style
abstract: |
  Text of abstract
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; Gobiidae

---

@@

```{r setup, echo = FALSE}    
# {r settings, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

@@

@@

# storing ALL sequences in a single fasta file
write(Sicydium_COI_seqs1, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n")
write(Sicydium_COI_seqs2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
write(Awaous_bananaCOI_seqs, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
write(Awaous_ocellarisCOI_seqs, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)

@@

Sicydium_COI_seqinrFormat <- read.fasta(file ="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

@@

# using library "ape"

Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run

#attr(Sicydium_COI_apeFormat, "species")
#names(Sicydium_COI_apeFormat)

@@

# writing a new fasta file (2nd) using the vector above 
write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10,
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat")

@@

### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
nameFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat",
                        seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them       # 562 seqs

# writing a new fasta (3rd) file with only those seqs with full scientific names
write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs],
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta")

### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
lengthFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta",
                          seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]       
write.fasta(sequences = fastaLength, names = names(fastaLength),
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta")  # 401 seqs

### Removing DUPLICATE sequences
# using "seqinr" to do it 
# for not using attributes to name sequences
duplicatesFasta <- read.fasta("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta",
                              set.attributes = FALSE) 
seqnames <- names(duplicatesFasta)
dups <- grep(TRUE,duplicated(duplicatesFasta))
uniqueNames <- seqnames[-dups]
uniqueSeqs <- duplicatesFasta[-dups] 
write.fasta(sequences = uniqueSeqs, names = uniqueNames,
            file.out = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")     # 204 seqs

# adding samples from Cuba (n=6)
#mySamples <- read.fasta("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_Cuba.fasta")
#write.fasta(mySamples, names(mySamples),
 #           "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta", open = "a")    # 210 seqs

@@

alignmentFasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")
ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")
#print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")

# REMOVING samples with to many GAPS at the beginning and the end of the sequences 
colmask(ClustalOAlignment) <- IRanges(start=c(1,666),end=c(25,679))                
ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
writeXStringSet(ClustalOAlignment2,
                file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta")   # 210 seqs -- 640 bp

### Replacing gaps at the begining and end with Ns
gapsFasta <- read.fasta(file = "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)
replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
write.fasta(replacingGaps, names(replacingGaps), "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")

@@

alignedFasta <- readDNAStringSet("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")

getRF <- function(alignedFasta){
  if (class(alignedFasta)!= c("DNAStringSet")){
    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
  }
  if (any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0){
    stop("This function do not take gaps as values/strings. Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
  }
  # predORF() do not read seqs with gaps   # predicts longest ORF for sense strand in each query sequence               
  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")                                       
  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
  codonPosition1 <- longestSeq1[1]/3
  codonConfirmation1 <- round(codonPosition1)
  length(alignedFasta)
  if (codonPosition1==codonConfirmation1){
   file.rename("/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
  print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
    } else {
    if (codonPosition1!=codonConfirmation1){
      tempLength <- length(alignedFasta[[1]])
      #remove first nucleotide position
      newRF2 <- subseq(alignedFasta, start = 2, end = tempLength)      # 639 bp
      # getting new RF from the new alignment with -1 bp.
      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense") 
      # checking size gain
      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
      codonPosition2 <- (longestSeq2[1]+2)/3
      codonConfirmation2 <- round(codonPosition2)
      if (codonPosition2==codonConfirmation2){
        writeXStringSet(newRF2, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
      print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
      } else {
      if (codonPosition2!=codonConfirmation2){
        #remove first nucleotide position
        tempLength <- length(alignedFasta[[1]])
        newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)
        writeXStringSet(newRF3, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
      print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
     }
   }  
  }
 }
}

getRF(alignedFasta)

@@

fastaRF <- read.dna(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")

@@

bestModel <- getBestModel(fastaRF)                # getting the AICc for each competing model

@@

fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY")     # seeking a maximum likelihood solution
#set.seed(123)    # getting the same tree topology each time the code runs 
bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE, mc.cores = 4)

# saving the tree in the folder 'analysis/figures/'
bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
optimalTree <- plotBS(bsTree, type = "p", cex = 0.2, font = 1)
# export tree with bootstrap values as node labels
write.tree(optimalTree, file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")

@@

(Figure 3.1)

@@

```{r inline-code2, include = TRUE, results='hide'}
# the result wont be displayed in the final document
loglikelihood <- round(-1*(as.numeric(fit[1])), 2)
loglikelihood

@@

```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus Sicydium (n=197). Sequences highlited in green represent samples from Cuba (n=6). Numbers on branches represent boostrap values higher than 70.", fig.width = 11.8, fig.height = 9.84)}

# Editing and visualizing the tree to represent --- library: ggtree
tree <- read.tree(file="/Users/sheila_ffish/Documents/rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")

@@

Sicydium_ML_tree <-suppressWarnings(suppressMessages(print(ggtree(tree, size = .4) + 
  xlim(-0.02, 0.26) +
  geom_treescale(x=0, y=-1.5, offset = -4) + 
  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
  geom_cladelabel(node=218, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=117, label="Sicydium brevifile", color="blue", hjust = -0.05, vjust = 0.3, fontsize = 3.5, fontface="italic") + 
  geom_cladelabel(node=320, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=328, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=379, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=415, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
  geom_cladelabel(node=212, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
  geom_hilight(node=377, fill="green") + 
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")))))

#Sicydium_ML_tree
```

```{r demo-plot, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

>>>>

---
title: 'The Truth is Revealed: Identity of the Fish Fry "Teti" Fishery in Eastern Cuba'
author:
  - Sheila Rodriguez-Machado:
      email: srodr62@lsu.edu
      institute: [UofO]
      correspondence: true
  - Tomas M. Rodriguez-Cabrera:
      email: tomasmichelrodriguez@gmail.com
      institute: [UofA]
      correspondence: false
institute:
  - UofO: Museum of Natural Sciences, Louisiana State University
  - UofA: Sociedad Cubana de Zoologia
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    bookdown::word_document2:
      fig_caption: yes
      reference_docx: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/template.docx' # Insert path for the DOCX file
      pandoc_args:
      - --lua-filter=../templates/scholarly-metadata.lua
      - --lua-filter=../templates/author-info-blocks.lua
      - --lua-filter=../templates/pagebreak.lua
bibliography: references.bib
csl: '/Users/sheila_ffish/Documents/rrcompendium/analysis/templates/journal-of-archaeological-science.csl' # Insert path for the bib-style
abstract: |
  texting texting
keywords: |
  amphidromy; Busck's Stone-Biting Goby; COI; Gobiidae

editor_options: 
  chunk_output_type: console
---

@@

```{r setup, echo = FALSE}    
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = FALSE,
  comment = "#>",
  fig.path = "../figures/",
  dpi = 300
)

@@

#getwd()

@@

# storing ALL sequences in a single fasta file
write(Sicydium_COI_seqs1, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n")
write(Sicydium_COI_seqs2, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)   # 'append = TRUE', I can keep adding seqs
write(Awaous_bananaCOI_seqs, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)
write(Awaous_ocellarisCOI_seqs, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", sep = "\n", append = TRUE)

@@

Sicydium_COI_seqinrFormat <- read.fasta(file ="../rrcompendium/analysis/data/raw_data/Sicydium_COI_1-seqs.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

@@

# using library "ape"
Sicydium_COI_apeFormat <- read.GenBank(Sicydium_COI_names)         # this line takes a few minutes to run

@@

# writing a new fasta file (2nd) using the vector above 
write.fasta(sequences = Sicydium_COI_seqinrFormat, names = Sicydium_COI_GenBank_Ids, nbchar = 10,
            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta")

@@

### REMOVING sequences of incomplete/unspecific NAMES; that is, those with 'sp.'
nameFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_2-seqinrFormat.fasta",
                        seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)

fastaNames <- strsplit(names(nameFasta), '_')                  # creating a vector splitting species names
completeSeqs <- which(! grepl("sp.", fastaNames))              # selecting only those with specific name in them

# writing a new fasta (3rd) file with only those seqs with full scientific names
write.fasta(sequences = nameFasta[completeSeqs], names = names(nameFasta)[completeSeqs],
            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta")

### REMOVING sequences of LENGTH shorter than 600 pb, given that the gene COI has between 627 and 658 bp.
lengthFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_3-names.fasta",
                          seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE)                     # 586 seqs
fastaLength <- lengthFasta[which(getLength(lengthFasta) >= 600)]
write.fasta(sequences = fastaLength, names = names(fastaLength),
            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta")             # 401 seqs

### Removing DUPLICATE sequences
# using "seqinr" to do it 
# for not using attributes to name sequences
duplicatesFasta <- read.fasta("../rrcompendium/analysis/data/raw_data/Sicydium_COI_4-lengths.fasta", set.attributes = FALSE)

seqnames <- names(duplicatesFasta)
dups <- grep(TRUE,duplicated(duplicatesFasta))
uniqueNames <- seqnames[-dups]
uniqueSeqs <- duplicatesFasta[-dups] 
write.fasta(sequences = uniqueSeqs, names = uniqueNames,
            file.out = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")               # 204 seqs

# adding samples from Cuba (n=6)
mySamples <- read.fasta("../rrcompendium/analysis/data/raw_data/Sicydium_COI_Cuba.fas")
write.fasta(mySamples, names(mySamples),
            "../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta", open = "a")              # 210 seqs

@@

alignmentFasta <- readDNAStringSet("../rrcompendium/analysis/data/raw_data/Sicydium_COI_5-dupli.fasta")
ClustalOAlignment <- msa(alignmentFasta, "ClustalOmega")
# print(ClustalOAlignment, showConsensus=FALSE, showNames=TRUE, show="complete")

# Checking for the position of sequences that were not edited properly so they still have nucleotides different that A, G, C, T 
nucType <- alphabetFrequency(ClustalOAlignment)
unknownNuc <- nucType[, c(5:14)]
unknownPosition <- which(unknownNuc > 0, arr.ind=T)

# REMOVING samples with too many GAPS at the beginning and the end of the sequences and UNKNOWN nucleotides (ex.: W, K, S) 
colmask(ClustalOAlignment) <- IRanges(start=c(1,666),end=c(25,679))
rowmask(ClustalOAlignment) <- IRanges(start=c(1),end=c(1))
ClustalOAlignment2 <- as(ClustalOAlignment, "DNAStringSet")
writeXStringSet(ClustalOAlignment2,
                file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta")           # 209 seqs -- 640 bp

### Replacing gaps at the begining and end with Ns
gapsFasta <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
#getLength(gapsFasta)
replacingGaps <- lapply(gapsFasta, function(x) gsub("-", "N", x))
write.fasta(replacingGaps, names(replacingGaps), "../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")

@@

alignedFasta <- readDNAStringSet("../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta")

getRF <- function(alignedFasta){
  if (class(alignedFasta)!= c("DNAStringSet")){
    stop("This function only takes files of class 'DNAStringSet'. Use package 'Biostrings' to read the input file")
  }
  if (any(letterFrequency(alignedFasta, "-", as.prob=TRUE))>0){
    stop("This function do not take gaps as values/strings. Go and replaced gaps with Ns (if the gaps are at the begining or end of the sequences) or choose a subset of sequences that do not contain gaps")
  }
  orf1 <- predORF(alignedFasta[1:10], n=1, type="df", mode="orf", strand="sense")   # predORF() do not read seqs with gaps
  # predicts longest ORF for sense strand in each query sequence                                                                          
  longestSeq1 <- orf1$start[orf1$width == max(orf1$width)]
  codonPosition1 <- longestSeq1[1]/3
  codonConfirmation1 <- round(codonPosition1)
  if (codonPosition1==codonConfirmation1){
   file.rename("../rrcompendium/analysis/data/raw_data/Sicydium_COI_7-gaps.fasta", "../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")                     # 640 bp
  print("The Reading Frame of these seqs is +1, no further action is required, go to modelTesting")
    } else {
    if (codonPosition1!=codonConfirmation1){
      tempLength <- length(alignedFasta[[1]])
      #remove first nucleotide position
      newRF2 <- subseq(alignedFasta, start = 2, end = tempLength)      # 639 bp
      # getting new RF from the new alignment with -1 bp.
      orf2 <- predORF(newRF2[1:10], n=1, type="df", mode="orf", strand="sense") 
      # checking size gain
      longestSeq2 <- orf2$start[orf2$width == max(orf2$width)]
      codonPosition2 <- (longestSeq2[1]+2)/3
      codonConfirmation2 <- round(codonPosition2)
      if (codonPosition2==codonConfirmation2){
        writeXStringSet(newRF2, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")
      print("The Reading Frame of these seqs was +2, new fasta file created, go to modelTesting")
      } else {
      if (codonPosition2!=codonConfirmation2){
        #remove first nucleotide position
        tempLength <- length(alignedFasta[[1]])
        newRF3 <- subseq(alignedFasta, start = 3, end = tempLength)
        writeXStringSet(newRF3, file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta")    # 638 bp
      print("The Reading Frame of these seqs was +3, new fasta file created, go to modelTesting")
     }
   }  
  }
 }
}

suppressWarnings(suppressMessages(print(getRF(alignedFasta))))

@@

fastaRF <- read.dna(file="../rrcompendium/analysis/data/raw_data/Sicydium_COI_newRF.fasta", format = "fasta")

fastaRF1 <- read.fasta(file = "../rrcompendium/analysis/data/raw_data/Sicydium_COI_6-alignment.fasta", seqtype = "DNA", as.string = TRUE, forceDNAtolower = FALSE, apply.mask = FALSE)
getLength(fastaRF1)

@@

bestModel <- suppressWarnings(suppressMessages(getBestModel(fastaRF)))               # getting the AICc for each competing model

@@

fit = optim.pml(fitStart, rearrangement = "stochastic", optGamma=TRUE, optInv=FALSE, model="HKY")     # seeking a maximum likelihood solution
set.seed(123)    # getting the same tree topology each time the code runs 
bs = bootstrap.pml(fit, bs=100, optNni=TRUE, multicore=TRUE, mc.cores = 4)

# saving the tree in the folder 'analysis/figures/'
bsTree <- plotBS(midpoint(fit$tree), bs, p = 75, type="p", bs.col = "blue")     # fit$tree is where the optimal tree is stored
optimalTree <- plotBS(bsTree, type = "p", cex = 0.2, font = 1)
# export tree with bootstrap values as node labels
write.tree(optimalTree, file="../rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")

@@

@@

```{r demo-plot1, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```

```{r inline-code2, include = TRUE, results='hide'}
# the result wont be displayed in the final document
loglikelihood <- round(-1*(as.numeric(fit[1])), 2)

@@

```{r Sicydium_ML_tree, fig.cap = paste("ML phylogenetic tree  (-lnL =", loglikelihood,") depicting the relationships among all available COI sequences of the genus *Sicydium*. Sequences highlited in green represent samples from Cuba. Numbers on branches represent boostrap values higher than 70.")}

# , fig.width = 11.8, fig.height = 10

# Editing and visualizing the tree to represent --- library: ggtree
tree <- read.tree(file="../rrcompendium/analysis/data/derived_data/Sicydium_ML_tree.tre")

# checkinf whether the tree is rooted
rootedTree <- is.rooted(tree)   # package ape. showing the tree root: geom_rootpoint(color="pink", size=5)

@@

Sicydium_ML_tree <-suppressWarnings(suppressMessages(print(ggtree(tree, size = .4) + 
  xlim(-0.02, 0.26) +
  geom_treescale(x=0, y=-1.5, offset = -4) + 
  geom_text2(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 80), size = 3, col= "black", vjust = -0.3, hjust=1.9, na.rm = FALSE) + 
  geom_nodepoint(aes(label=label, subset = !is.na(as.numeric(label)) & as.numeric(label) > 70), size = 1, col= "black") + 
  geom_cladelabel(node=216, label="Sicydium plumieri", color="red", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=108, label="Sicydium brevifile", color="blue", hjust = -0.01, vjust = 0.2, fontsize = 3.5, fontface="italic") + 
  geom_cladelabel(node=408, label="Sicydium salvini", color="purple", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=356, label="Sicydium punctatum", color="brown", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=320, label="Sicydium buscki", color="black", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") + 
  geom_cladelabel(node=407, label="Sicydium crenilabrum", color="orange", hjust = -0.03, vjust = 0.3, fontsize = 3.5, offset = 0.004, fontface="italic") +
  geom_cladelabel(node=211, label="Awaous spp.", color="darkgreen", hjust = -0.07, fontsize = 3.5, offset = 0.002, fontface="italic") +
  geom_hilight(node=336, fill="green") +
  #geom_hilight(node=332, fill="green") +
  geom_hilight(node=135, fill="green") +
  geom_hilight(node=121, fill="green") +
  geom_hilight(node=120, fill="green") +
  geom_hilight(node=139, fill="green") +
  theme(plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")))))

```{r demo-plot2, fig.cap="A plot of random numbers"}
plot(rnorm(10))
```
